---
title: "R Notebook"
output: html_notebook
---

```{r}
library(devtools)
install_github("riv","tomasgreif")
```



```{r}
library(tidyverse)
library(Information)
library(ggplot2)
```


```{r}
options(digits=2)

```


# Motivation


#In credit score context, defaults needs to be avoided, the default is a "bad". Therefore negative WoE are good, because it tells that the probability of event (loan) is unlikely to happen compared to the non-event (not default)
# The WoE is used as a continuous variable in the model, and has linear relationship with the log-odds of the dependent variable
# Models using continuous variables often perform better and are easier to interpret than those using raw categorical variables.


```{r}
loan <- read_csv("data/cr_loan2.csv")
head(loan)
```






```{r}
calculate <- function(data, feature, target) {
  dist_table <- data |> 
    summarise(good = sum(!!sym(target) == 1),
              bad = sum(!!sym(target) == 0),
              .by = loan_intent) |> 
    mutate(total = good+bad,
           dist_good = good/sum(good),
           dist_bad = bad/sum(bad)) |> 
    filter(total > 0)
  
  woe_table <- dist_table |>
    mutate(
      proportion = dist_bad/dist_good,
      woe = log(dist_bad/dist_good))
  return(woe_table)
    }

woe_table <- calculate(loan, "loan_intent", "loan_status")
woe_table |> head()
```


```{r}
calculate <- function(data, feature, target) {
  dist_table <- data |> 
    summarise(bad = sum(!!sym(target) == 1),
              good = sum(!!sym(target) == 0),
              total = n(),
              .by = loan_intent) |> 
    mutate(total = good+bad,
           percentage_good = good/sum(good),
           percentage_bad = bad/sum(bad),
           woe = log(percentage_good/percentage_bad),
           iv_parc = woe * (percentage_good - percentage_bad))
  }


woe_table <- calculate(loan, "loan_intent", "loan_status")
woe_table |> head()
woe_table |> summarise(sum(iv_parc))
```


```{r}
calculate_iv <- function(data, target, variable) {
  df <- data %>% 
    group_by(!!sym(variable)) %>% 
    summarise(Count = n(),
              Good = sum(!!sym(target) == 0),
              Bad = sum(!!sym(target) == 1)) %>%
    mutate(Distr_Good = Good / sum(Good),
           Distr_Bad = Bad / sum(Bad),
           WoE = log(Distr_Good / Distr_Bad),
           IV = (Distr_Good - Distr_Bad) * WoE) %>%
    ungroup()
  
  total_iv <- sum(df$IV)
  return(list(IV_table = df, Total_IV = total_iv))}

calculate_iv(loan, "loan_status", "loan_intent")

```

